---
title: "Hyderabad Annual Tree Survey Report"
author: "Ram Dayal Vaishnav"
date: "`r Sys.Date()`"
output: html_document
#output: pdf_document
---

## Hyderabad Annual Tree Survey Report

With rapid urbanization transforming Indian cities, understanding and preserving urban biodiversity is crucial. The Hyderabad Annual Tree Survey leverages citizen science to document and conserve the city's flora and fauna.

Methodologically, the survey employed a custom web app and iNaturalist for data collection. Volunteers, organized into teams, surveyed both avenue and non-avenue locations, recording detailed information on tree species, diameter, and related observations, alongside animal counts and activities.

A pilot survey refined these methods, leading to a comprehensive survey of over 750 trees and 120 species across Hyderabad. This initiative provides essential insights into species distribution and the impacts of urbanization on biodiversity, guiding future conservation efforts.

## Data Analysis

### Tree Species and Observations

The survey covered over 750 trees, identifying 120 species. Key findings include:

```{r include=F}
# Load necessary libraries
library(ggplot2)
library(dplyr)

treeData  <- read.csv("data/TreeData.csv")

# Assuming treeData is your data frame
tree_counts <- treeData %>%
  group_by(TreeSpeciesName) %>%
  summarize(TreeCount = n()) %>%
  arrange(desc(TreeCount))

# Remove the "Unidentified / Other" entry
filtered_tree_counts <- tree_counts %>%
  filter(TreeSpeciesName != "Unidentified / Other")

# Get the top 10 species after removing "Unidentified / Other"
top_10_filtered_tree_counts <- filtered_tree_counts %>%
  top_n(10, wt = TreeCount)
```

```{r out.width="60%", echo=F}
# Create the bar chart
ggplot(top_10_filtered_tree_counts, aes(x = reorder(TreeSpeciesName, TreeCount), y = TreeCount)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Top 10 Tree Species by Count",
       x = "",
       y = "") +
  theme_minimal()

```

```{r include=F}
nativeness <- filtered_tree_counts %>%
  left_join(treeData %>% select(TreeSpeciesName, Nativeness) %>% distinct(), by = "TreeSpeciesName")

# Summarize the data by Nativeness
nativeness_summary <- nativeness %>%
  group_by(Nativeness) %>%
  summarize(
    TotalTreeCount = sum(TreeCount),
    SpeciesCount = n(),
    ExampleSpecies = paste(unique(TreeSpeciesName), collapse = ", ")
  ) %>%
  arrange(desc(TotalTreeCount))

# Print the summarized data nicely
print(nativeness_summary)
```

```{r out.width="60%", echo=F}
# Create a bar chart to display the total tree count by nativeness with species names
ggplot(nativeness_summary, aes(x = reorder(Nativeness, SpeciesCount), y = SpeciesCount, fill = Nativeness)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = ExampleSpecies),nudge_x = -0.2, hjust = 0.1, size = 3.5, color = "black") +  # Add species names
  #coord_flip() +  # Flip coordinates for better readability
  labs(title = "Species by Nativeness",
       x = "Nativeness",
       y = "Species Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")  + # Optional: use a color palette for better visualization
  theme(legend.position = "none")  # Remove the legend


```

```{r include=F}
# install.packages("tidyverse")
# install.packages("patchwork")

library(patchwork)
library(tidyverse)
library(dplyr)

treeData  <- read.csv("data/TreeData.csv")
animalData <- read.csv("data/AnimalData.csv")
treeAnimalData  <- read.csv("data/TreeAnimalData.csv")

colnames(treeAnimalData)
colnames(treeData)

identified_significant_tree_counts <- treeData %>%
  group_by(TreeSpeciesName) %>%
  summarize(TreeCount = n()) %>%
  arrange(desc(TreeCount)) %>%
  filter(TreeSpeciesName != "Unidentified / Other" & TreeCount > 5) 

nativeness_summary_significant <- identified_significant_tree_counts%>%
  left_join(treeData %>% select(TreeSpeciesName, Nativeness) %>% distinct(), by = "TreeSpeciesName") %>%
  group_by(Nativeness) %>%
  summarize(SpeciesCount = n(), AllTreesCount = sum(TreeCount)) %>%
  # Add a total row
  bind_rows(
    tibble(
      Nativeness = "Grand Total",
      SpeciesCount = sum(.$SpeciesCount),
      AllTreesCount = sum(.$AllTreesCount)
    )
  )

nativeness_summary_significant

# Nativeness-Animal Relation ####

# Filter treeAnimalData to include only significant tree species and join with tree counts
animal_summary_table <- treeAnimalData %>%
  filter(TreeSpeciesName %in% identified_significant_tree_counts$TreeSpeciesName) %>%
  group_by(Nativeness) %>%
  summarize(
    SUM_TotalAnimals = sum(AnimalCount, na.rm = TRUE),
  ) %>%
  # Add a total row
  bind_rows(
    tibble(
      Nativeness = "Grand Total",
      SUM_TotalAnimals = sum(.$SUM_TotalAnimals)
    )
  )

animal_summary_table

combined_summary <- nativeness_summary_significant %>%
  left_join(animal_summary_table, by = "Nativeness") %>%
  mutate(
    AVG_AnimalsPerTree = ifelse(AllTreesCount > 0, SUM_TotalAnimals / AllTreesCount, NA)  # Calculate average animals per tree
  )

# View the combined summary
print(combined_summary)




# Create a data frame for plotting to handle both TreeCount and TotalAnimals together
plot_data <- combined_summary %>%
  select(Nativeness, AllTreesCount, SUM_TotalAnimals, AVG_AnimalsPerTree) %>%
  pivot_longer(cols = c(AllTreesCount, SUM_TotalAnimals, AVG_AnimalsPerTree),
               names_to = "Metric",
               values_to = "Value")

# Plot Tree Count and Total Animals using bar chart
tree_animals_plot <- ggplot(plot_data %>% filter(Metric != "AVG_AnimalsPerTree"),
                            aes(x = Nativeness, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Tree Count and Total Animals by Nativeness",
       x = "Nativeness",
       y = "Count",
       fill = "Metric") +
  theme_minimal() +
  coord_flip() +
  scale_fill_manual(values = c("AllTreesCount" = "#4CAF50", "SUM_TotalAnimals" = "orange"),
                    labels = c("Trees", "Animals"))  # Custom labels

# Plot Average Animals per Tree using line plot
avg_animals_plot <- ggplot(plot_data %>% filter(Metric == "AVG_AnimalsPerTree"),
                           aes(x = Nativeness, y = Value, group = 1)) +
  geom_line(color = "purple") +
  geom_point(color = "purple") +
  labs(title = "Average Animals per Tree by Nativeness",
       x = "Nativeness",
       y = "Average Animals per Tree") +
  theme_minimal() +
  coord_flip()

# Combine plots
combined_plot <- tree_animals_plot / avg_animals_plot
```

```{r out.width="60%", echo=F}
# Display the combined plot
print(combined_plot)
```

```{r include=F}
# Data for Avenue vs. Non-Avenue Trees
avenue_non_avenue_data <- data.frame(
  Type = c("Avenue Trees", "Non-Avenue Trees"),
  PercentageTreeCount = c(52.32, 41.04),
  PercentageAnimals = c(70.39, 26.99),
  AvgAnimalsPerTree = c(10.8, 5.3)
)
library(ggplot2)
library(tidyr)

# Reshape the data for plotting
avenue_non_avenue_long <- avenue_non_avenue_data %>%
  pivot_longer(cols = c(PercentageTreeCount, PercentageAnimals, AvgAnimalsPerTree),
               names_to = "Metric",
               values_to = "Value")

# Plot Avenue vs. Non-Avenue Trees
avenue_non_avenue_plot <- ggplot(avenue_non_avenue_long, aes(x = Type, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Avenue vs. Non-Avenue Trees",
       x = "Tree Type",
       y = "Value",
       fill = "Metric") +
  theme_minimal() +
  scale_fill_manual(values = c("PercentageTreeCount" = "#1f77b4", 
                               "PercentageAnimals" = "#ff7f0e",
                               "AvgAnimalsPerTree" = "#2ca02c"),
                    labels = c("Percentage of Tree Count", "Percentage of Animals", "Avg. Animals per Tree"))


```

```{r out.width="60%", echo=F}

print(avenue_non_avenue_plot)

```

```{r include=F}
treeData  <- read.csv("data/TreeData.csv")
animalData <- read.csv("data/AnimalData.csv")
treeAnimalData  <- read.csv("data/TreeAnimalData.csv")

aqi_summary <- treeAnimalData %>%
  filter(AQI != '-') %>%
  group_by(AQI) %>%
  summarize(
    Total_Trees = n(),
    Total_Animals = sum(AnimalCount, na.rm = TRUE),
    Average_Animals_Per_Tree = mean(AnimalCount, na.rm = TRUE)
  )

library(ggplot2)
```

```{r out.width="60%", echo=F}

# Plot the summary statistics
ggplot(aqi_summary, aes(x = factor(AQI, levels = rev(unique(AQI))), y = Average_Animals_Per_Tree, fill = AQI)) +
  geom_bar(stat = "identity") +
  labs(title = "Impact of AQI on Average Animals per Tree",
       x = "AQI Category",
       y = "Average Animals per Tree") +
  theme_minimal() + 
  theme(legend.position = 'none')


```

## Recommendations

1.  **Prioritize Native Species:** Native Indian trees support significantly more biodiversity compared to introduced species.
2.  **Enhance Urban Green Spaces:** Avenue trees play a crucial role in supporting urban wildlife.
3.  **Improve Air Quality:** Higher air quality correlates with greater biodiversity.

## Conclusion

The Hyderabad Annual Tree Survey underscores the value of citizen science in urban biodiversity research. Future surveys should address vehicle density, human activity, and seasonal variations to further understand urban ecosystems.

------------------------------------------------------------------------

**References:** - [HAT Survey iNaturalist Project](#)
